---
title: "ClusteringPigBAL"
author: "Arianne Richard"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

This script picks up where "processing all.Rmd" left off.

We will load the data, cluster, remove doublet clusters, make direct cluster comparisons, etc, in order to come up with the best dataset for annotation.

# Clustering

## Reading in the data

```{r}
library(SingleCellExperiment)
library(DropletUtils)
library(scran)
library(scater)
library(uwot)
library(biomaRt)
library(scds)
library(batchelor)

load('ongoing_full.RData')

```

## Clustering the data

Now we cluster the data, look at sample contribution to each cluster, and find marker genes for each cluster for annotation.

We'll use the graph-based clustering implemented in bluster, specifically using jaccard index to weight edges and the louvain method for community detection, with a k value of 20. This is largely because this combination gave us a reasonable number of clusters with decent quality metrics after trying a variety of different options. Thus we will run this on the batch-corrected data and look at distribution of samples across clusters and top marker genes for each cluster.

```{r}
library(bluster)

set.seed(200)
sce_all_corrected_clust <- clusterCells(sce_all_corrected,
                                        use.dimred="corrected",
                                        BLUSPARAM=SNNGraphParam(k=20, 
                                                                type="jaccard",
                                                                cluster.fun="louvain"))

table(sce_all_corrected_clust)
sce_all_corrected$cluster <- sce_all_corrected_clust

plotUMAP(sce_all_corrected, colour_by="cluster", point_size=0.3)
plotTSNE(sce_all_corrected, colour_by="cluster", point_size=0.3)
plotReducedDim(sce_all_corrected, dimred='corrected', colour_by="cluster", point_size=0.1, ncomponents=5)


dat <- as.data.frame(table(sce_all_corrected$cluster, sce_all_corrected$Sample))
colnames(dat) <- c('cluster', 'sample', 'n_cells')

ggplot(dat, aes(fill=sample, y=n_cells, x=cluster)) +
  geom_bar(position='stack', stat='identity')

## check columns match before adding count data to sce_all_corrected
table(colnames(sce_all_corrected) == colnames(sce_all))

assays(sce_all_corrected)$counts <- assays(sce_all)$counts[rownames(sce_all_corrected),]
assays(sce_all_corrected)$logcounts <- assays(sce_all)$logcounts[rownames(sce_all_corrected),]

marker.info.perbatch <- scoreMarkers(sce_all_corrected, 
                                     groups=sce_all_corrected$cluster,
                                     block=sce_all_corrected$run,
                                     assay.type='logcounts',
                                     full.stats=TRUE)

for(i in 1:length(marker.info.perbatch)){
  ord <- marker.info.perbatch[[i]][order(marker.info.perbatch[[i]]$mean.logFC.cohen, decreasing=TRUE),]
  ord$symbol <- rowData(sce_all)[rownames(ord), 'Symbol']
  write.table(ord, 
              paste0('Results/cluster', names(marker.info.perbatch)[i], 'markers.txt'), 
              quote=F, sep='\t')
}

```

## Evaluating the clustering 

We will now run several metrics to evaluate how well our cells are clustered, thinking about whether they have been either over- or under-clustered. 

First we'll look at the silhouette widths, which incorporates both the distances to other cells in the same cluster as well as the distances to cells in other clusters. The plot we make will colour cells according to their nearest cluster if they have a negative silhouette width.

```{r}

sil.approx <- approxSilhouette(reducedDim(sce_all_corrected, "corrected"), clusters=sce_all_corrected$cluster)
sil.approx
sil.data <- as.data.frame(sil.approx)
sil.data$closest <- factor(ifelse(sil.data$width > 0, sce_all_corrected$cluster, sil.data$other))
sil.data$cluster <- sce_all_corrected$cluster

ggplot(sil.data, aes(x=cluster, y=width, colour=closest)) +
    ggbeeswarm::geom_quasirandom(method="smiley")
```

This shows some potential issues particularly with clusters 5 and 17.

To better separate the factors that contribute to this metric, we'll divide out purity and heterogeneity, first looking at cluster purity. Here again cells are coloured according to their nearest cluster if their purity is particularly low.

```{r}

pure <- neighborPurity(reducedDim(sce_all_corrected, "corrected"), sce_all_corrected$cluster)
pure

pure.data <- as.data.frame(pure)
pure.data$maximum <- factor(pure.data$maximum)
pure.data$cluster <- sce_all_corrected$cluster

ggplot(pure.data, aes(x=cluster, y=purity, colour=maximum)) +
    ggbeeswarm::geom_quasirandom(method="smiley")

table(Cluster=sce_all_corrected$cluster, pure.data$maximum)
```

Here it looks like the lowest purity is in clusters 1, 4, 6, and 7. 17 is somewhat unpure, which may have contributed to the silhouette width results, while cluster 5 looks extremely pure, suggesting its issue is more likely intra-cluster heterogeneity.

And finally we'll look at the heterogeneity within each cluster by computing the within cluster sum of squares.

```{r}
rmsd <- clusterRMSD(reducedDim(sce_all_corrected, "corrected"), sce_all_corrected$cluster)
barplot(rmsd, ylab="RMSD", xlab="Cluster")

```

And now we can see what the particularly low silhouette width from cluster 5 was all about - it's more heterogeneous than the others. 

Let's have a look at this cluster, also thinking about QC metrics.

```{r}
plotUMAP(sce_all_corrected[,sce_all_corrected$cluster=='5'], colour_by="batch", point_size=1)
plotTSNE(sce_all_corrected[,sce_all_corrected$cluster=='5'], colour_by="batch", point_size=1)
plotReducedDim(sce_all_corrected[,sce_all_corrected$cluster=='5'], dimred='corrected', colour_by="batch", point_size=0.1, ncomponents=5)

QC.tab <- data.frame(colData(sce_all_corrected)[,c('cluster', 'sum_counts', 'mito_ratio', 'genes_detected')])

ggplot(QC.tab, aes(x=cluster, y=log2(sum_counts))) +
    ggbeeswarm::geom_quasirandom(method="smiley")

ggplot(QC.tab, aes(x=cluster, y=log2(genes_detected))) +
    ggbeeswarm::geom_quasirandom(method="smiley")

ggplot(QC.tab, aes(x=cluster, y=mito_ratio)) +
    ggbeeswarm::geom_quasirandom(method="smiley")

props5 <- prop.table(table(sce_all_corrected$cluster, sce_all_corrected$Sample), margin=2)['5',]
barplot(props5)
dfprops5 <- annot[,c('Condition', 'Sex')]
dfprops5$propC5 <- props5[rownames(dfprops5)]
ggplot(dfprops5, aes(x=Condition, y=propC5)) +
  geom_dotplot(
  binaxis = "y", stackdir = "center",
  fill = "blue"
  )

```

So it looks from the UMAP and TSNE plots like there are some cells that are quite different than the bulk of the cluster for each of these. One option we have is to sub-cluster even more. It's possible the distant cells are just contaminating from other clusters. 

However, the other issue that becomes quite apparent when looking at the QC metrics is that the cells have substantially lower total counts and genes detected than the other clusters. From the marker genes, these cells don't really have specific positive defining features and their main shared characteristic seems to be lack of expression of particular genes. I suspect that this is a set of heterogeneous poor-quality cells that have been clustered together based on their dying features. The percentage in each sample varies but there don't appear to be any strong associations with condition. To me, the best bet will be exclusion.


## Doublet inspection

It may be the case that our suspected doublets, which fell in a particular region of the UMAP, have formed their own cluster. This is a convenient means of removing them as we can get rid of the whole cluster and don't have to rationalise a threshold for suspected doublet value. 

```{r}
prop.table(table(sce_all_corrected$cluster, sce_all_corrected$query_doublet), margin=1)

doublet.tab <- data.frame(colData(sce_all_corrected)[,c('cxds_score', 'cluster', 'query_doublet', 'sum_counts')])

ggplot(doublet.tab, aes(x=cluster, y=cxds_score, colour=query_doublet)) +
    ggbeeswarm::geom_quasirandom(method="smiley")

ggplot(doublet.tab, aes(x=cluster, y=sum_counts, colour=query_doublet)) +
    ggbeeswarm::geom_quasirandom(method="smiley")
```

These data strongly suggest that cluster 17 is largely composed of doublets. Let's look at the top marker genes to make sure we aren't getting rid of something unique, and check its representation across samples.

```{r}
plotUMAP(sce_all_corrected, colour_by='cluster')
plotUMAP(sce_all_corrected, by_exprs_values='logcounts', colour_by='ENSSSCG00000023333')
plotUMAP(sce_all_corrected, by_exprs_values='logcounts', colour_by='ENSSSCG00000000492')
plotUMAP(sce_all_corrected, by_exprs_values='logcounts', colour_by='ENSSSCG00000022230')

barplot(prop.table(table(sce_all_corrected$cluster, sce_all_corrected$Sample), margin=2)['17',])

```

This looks fine. On these grounds we should likely exclude cluster 17 as doublets.

The last thing to think about is cluster 11, which has a large percentage of suspected doublets and did exhibit a strange mix of B cell and myeloid markers. Let's look at these cells individually to see if there is a subset of co-expressers that could be some sort of doublets.

```{r}
plotUMAP(sce_all_corrected[,sce_all_corrected$cluster=='11'], colour_by="batch", point_size=1)
plotUMAP(sce_all_corrected[,sce_all_corrected$cluster=='11'], colour_by="query_doublet", point_size=1)

## CD163
CD163 <- rownames(sce_all_corrected)[rowData(sce_all_corrected)$Symbol == 'CD163']
plotUMAP(sce_all_corrected[,sce_all_corrected$cluster=='11'], colour_by=CD163, point_size=1, by_exprs_values='logcounts')
## JCHAIN
JCHAIN <- rownames(sce_all_corrected)[rowData(sce_all_corrected)$Symbol == 'JCHAIN']
plotUMAP(sce_all_corrected[,sce_all_corrected$cluster=='11'], colour_by=JCHAIN, point_size=1, by_exprs_values='logcounts')

### make some heatmaps of macrophages, B cells, and cluster 11, selecting marker genes from each subset to examine within cluster 11

library(pheatmap)

gene_list <- rownames(sce_all_corrected)[order(marker.info.perbatch$`11`$mean.logFC.cohen, decreasing=TRUE)][1:20]
# macrophages
gene_list <- c(gene_list, rownames(sce_all_corrected)[order(marker.info.perbatch$`11`$full.logFC.cohen$`1`, decreasing=TRUE)][1:20])
gene_list <- c(gene_list, rownames(sce_all_corrected)[order(marker.info.perbatch$`11`$full.logFC.cohen$`4`, decreasing=TRUE)][1:20])

gene_list <- gene_list[!duplicated(gene_list)]

mat <- exprs(sce_all_corrected)[gene_list, sce_all_corrected$cluster %in% c('1', '4')]
mat_clust <- sce_all_corrected$cluster[sce_all_corrected$cluster %in% c('1', '4')]
## downsample 
set.seed(100)
inds <- sample.int(length(mat_clust), 0.005*length(mat_clust), replace=FALSE)
mat <- mat[, inds]
mat_clust <- mat_clust[inds]

# B cells
mat2 <- exprs(sce_all_corrected)[gene_list, sce_all_corrected$cluster %in% c('2', '3')]
mat2_clust <- sce_all_corrected$cluster[sce_all_corrected$cluster %in% c('2', '3')]
## downsample 
set.seed(100)
inds <- sample.int(length(mat2_clust), 0.05*length(mat2_clust), replace=FALSE)
mat2 <- mat2[, inds]
mat2_clust <- mat2_clust[inds]

mat <- cbind(mat, mat2)
mat_clust <- c(mat_clust, mat2_clust)

mat <- cbind(mat, exprs(sce_all)[gene_list, sce_all_corrected$cluster == '11'])
mat_clust <- c(mat_clust, sce_all_corrected$cluster[sce_all_corrected$cluster =='11'])
rownames(mat) <- rowData(sce_all_corrected)[rownames(mat), 'Symbol']
colannot <- data.frame(cluster=paste0('c', as.character(mat_clust)))
rownames(colannot) <- colnames(mat)

pheatmap(mat, show_colnames=FALSE, cex=0.8, scale='row', annotation_col=colannot, border_color=NA, annotation_colors=list(cluster=c(c1='red', c2='blue', c3='darkgreen', c4='orange', c11='grey')))

fewgenes <- rownames(sce_all_corrected)[rowData(sce_all_corrected)$Symbol %in% c('CD163', 'JCHAIN')]
plotExpression(sce_all_corrected, features=fewgenes, x='cluster')
plotExpression(sce_all_corrected, features=fewgenes, x='cluster', exprs_values='counts')

```

Interestingly, the doublet annotation doesn't line up with the suspected doublets here. There are a handful of cells expressing a number of B cell markers and all of the cells express a pretty high level of JCHAIN. The cells also exhibit robust expression of cluster-defining myeloid genes (e.g. CD163). Thus, it is either a set of odd cells or doublets. It seems unlikely to be pure cluster contamination.

Based on this and discussion among ourselves, we have decided to define this cluster as doublets.


## Subclustering

Upon closer inspection and attempted annotation, we discovered that there are a couple of clusters that might benefit from subclustering. Specifically, clusters 14, 18 and 20 are composed of a mixture of CD4 and CD8 T cells, and to run differential abundance analyses according to sorted cell types we'll need to separate these. In addition, cluster 9 appears to be a heterogeneous mixture of cells expressing CD3 (T cells) and cells not expressing CD3, suggesting that it may benefit from subclustering.

In order to run subclustering, we start by subsetting cells in the cluster of interest. In an ideal world, CD4 and CD8 T cell-associated genes would dominate the highly variable genes and enable unsupervised sub-clustering to separate the cell types. Unfortunately, this is often not the case. We therefore need to supervise the clustering. To this end, we pull out CD4 and CD8B genes that encode the markers used for sorting and use a zero-inflated Kendall's Tau correlation metric implemented in SCHOT to identify the top genes among the highly variable genes correlated with these markers. We then use this filtered list for subclustering.

We'll start with cluster 14.

```{r}

c14 <- sce_all_corrected$cluster == '14'

CD4 <- counts(sce_all)['ENSSSCG00000000687',c14]
CD8 <- counts(sce_all)['ENSSSCG00000008213',c14]

library(scHOT)
varmod14 <- modelGeneVar(sce_all[,c14], block=sce_all[,c14]$run)
vargenes14 <- getTopHVGs(varmod14, n=1000)

CD4corr <- vector('numeric', length(vargenes14))
CD8corr <- CD4corr
for(i in 1:length(vargenes14)){
  CD4corr[i] <- weightedZIKendall(exprs(sce_all)['ENSSSCG00000000687', c14],
                               exprs(sce_all)[vargenes14[i], c14]
  )
  CD8corr[i] <- weightedZIKendall(exprs(sce_all)['ENSSSCG00000008213', c14],
                               exprs(sce_all)[vargenes14[i], c14]
  )
}

top <- vargenes14[c(which(CD4corr > 0.07),which(CD8corr > 0.07))]
top <- c(top, 'ENSSSCG00000000687', 'ENSSSCG00000008213')
top <- top[!duplicated(top)]

set.seed(100)
sce_all_14 <- fastMNN(sce_all[,c14], batch=sce_all[,c14]$run, subset.row = top, get.variance = TRUE, k=10)

set.seed(200)
sce_all_14_clust <- clusterCells(sce_all_14,
                                        use.dimred="corrected",
                                        BLUSPARAM=SNNGraphParam(k=10, 
                                                                type="jaccard",
                                                                cluster.fun="fast_greedy"))

table(sce_all_14_clust)
sce_all_14$cluster <- sce_all_14_clust
assays(sce_all_14)$logcounts <- assays(sce_all)$logcounts[rownames(sce_all_14), c14]

CD4CD8 <- rownames(sce_all)[rowData(sce_all)$Symbol %in% c('CD4', 'CD8B')]
plotExpression(sce_all_14, features=CD4CD8, x='cluster', exprs_values='logcounts')

```

Unfortunately, cluster 14 does not subcluster well, even in a supervised manner, into  CD4+ and CD8+ groups. We've tried several different mixtures of parameters and it just doesn't split. In fact, expression of CD4 and CD8 is more of a continuum in this cluster instead of bimodal as well. 

Let's eyeball whether there are different proportions of cells in this cluster by condition.

```{r}

props <- prop.table(table(sce_all_corrected$cluster, sce_all_corrected$Sample), margin=2)['14',]
proptab <- annot[,c('sample', 'Condition', 'Sex')]
proptab$props <- props[rownames(proptab)]

ggplot(data=proptab, aes(x=Condition, y=props)) +
  geom_boxplot() +
  geom_point(aes(col=Sex), position='jitter')


```

So this doesn't appear to be associated with condition. (Two of the pH1N1 values are very low, but one is high, and they are not segregated by sex.)  We may need to just leave this cluster out of our DA analysis groupings.

Next we do Cluster 18.

```{r}

c18 <- sce_all_corrected$cluster == '18'

CD4 <- counts(sce_all)['ENSSSCG00000000687',c18]
CD8 <- counts(sce_all)['ENSSSCG00000008213',c18]


library(scHOT)
varmod18 <- modelGeneVar(sce_all[,c18], block=sce_all[,c18]$run)
vargenes18 <- getTopHVGs(varmod18, n=1000)

CD4corr <- vector('numeric', length(vargenes18))
CD8corr <- CD4corr
for(i in 1:length(vargenes18)){
  CD4corr[i] <- weightedZIKendall(exprs(sce_all)['ENSSSCG00000000687', c18],
                               exprs(sce_all)[vargenes18[i], c18]
  )
  CD8corr[i] <- weightedZIKendall(exprs(sce_all)['ENSSSCG00000008213', c18],
                               exprs(sce_all)[vargenes18[i], c18]
  )
}

top <- vargenes18[c(which(CD4corr > 0.1),which(CD8corr > 0.1))]
top <- c(top, 'ENSSSCG00000000687', 'ENSSSCG00000008213')
top <- top[!duplicated(top)]

set.seed(100)
sce_all_18 <- fastMNN(sce_all[,c18], batch=sce_all[,c18]$run, subset.row = top, get.variance = TRUE)

set.seed(200)
sce_all_18_clust <- clusterCells(sce_all_18,
                                        use.dimred="corrected",
                                        BLUSPARAM=SNNGraphParam(k=20, 
                                                                type="jaccard",
                                                                cluster.fun="fast_greedy"))

table(sce_all_18_clust)
sce_all_18$cluster <- sce_all_18_clust
assays(sce_all_18)$logcounts <- assays(sce_all)$logcounts[rownames(sce_all_18), c18]

CD4CD8 <- rownames(sce_all)[rowData(sce_all)$Symbol %in% c('CD4', 'CD8B')]
plotExpression(sce_all_18, features=CD4CD8, x='cluster', exprs_values='logcounts')
```

This subclusters nicely. So let's change cluster 18 labels to 18a and 18b to reflect the subsetting. 

```{r}
sce_all_18_clust <- as.character(sce_all_18_clust)
sce_all_18_clust[sce_all_18_clust == '1'] <- '18a'
sce_all_18_clust[sce_all_18_clust == '2'] <- '18b'
cs <- as.character(sce_all_corrected$cluster)
cs[c18] <- sce_all_18_clust
sce_all_corrected$cluster <- cs
```

And cluster 20. Note that here we reduce the k nearest neighbours examined because the number of cells is so low.

```{r}

c20 <- sce_all_corrected$cluster == '20'

CD4 <- counts(sce_all)['ENSSSCG00000000687',c20]
CD8 <- counts(sce_all)['ENSSSCG00000008213',c20]


library(scHOT)
varmod20 <- modelGeneVar(sce_all[,c20], block=sce_all[,c20]$run)
vargenes20 <- getTopHVGs(varmod20, n=1000)

CD4corr <- vector('numeric', length(vargenes20))
CD8corr <- CD4corr
for(i in 1:length(vargenes20)){
  CD4corr[i] <- weightedZIKendall(exprs(sce_all)['ENSSSCG00000000687', c20],
                               exprs(sce_all)[vargenes20[i], c20]
  )
  CD8corr[i] <- weightedZIKendall(exprs(sce_all)['ENSSSCG00000008213', c20],
                               exprs(sce_all)[vargenes20[i], c20]
  )
}

top <- vargenes20[c(which(CD4corr > 0.10),which(CD8corr > 0.10))]
top <- c(top, 'ENSSSCG00000000687', 'ENSSSCG00000008213')
top <- top[!duplicated(top)]

set.seed(100)
sce_all_20 <- fastMNN(sce_all[,c20], batch=sce_all[,c20]$run, subset.row = top, get.variance = TRUE, k=9)

set.seed(200)
sce_all_20_clust <- clusterCells(sce_all_20,
                                        use.dimred="corrected",
                                        BLUSPARAM=SNNGraphParam(k=9, 
                                                                type="jaccard",
                                                                cluster.fun="fast_greedy"))

table(sce_all_20_clust)
sce_all_20$cluster <- sce_all_20_clust
assays(sce_all_20)$logcounts <- assays(sce_all)$logcounts[rownames(sce_all_20), c20]

CD4CD8 <- rownames(sce_all)[rowData(sce_all)$Symbol %in% c('CD4', 'CD8B')]
plotExpression(sce_all_20, features=CD4CD8, x='cluster', exprs_values='logcounts')
```

This clustering did a decent job except for 3 rogue cells that appear to be in the wrong cluster. They must have some similarities to the others. We'll go with this subclustering and change the labels to 20a and 20b.

```{r}

sce_all_20_clust <- as.character(sce_all_20_clust)
sce_all_20_clust[sce_all_20_clust == '1'] <- '20a'
sce_all_20_clust[sce_all_20_clust == '2'] <- '20b'
cs <- as.character(sce_all_corrected$cluster)
cs[c20] <- sce_all_20_clust
sce_all_corrected$cluster <- cs

```

Finally we would like to subcluster cluster 9 as it looked as though this was made of 2 mutually exclusive cell types, one expressing CD3E/CD8A and the other with more KLRB1 and RORC. Here we will just be guided by the data and go for an unsupervised split to see what we get.

```{r}
c9 <- sce_all_corrected$cluster == '9'
varmod9 <- modelGeneVar(sce_all[,c9], block=sce_all[,c9]$run)
vargenes9 <- getTopHVGs(varmod9, n=2000)

set.seed(100)
sce_all_9 <- fastMNN(sce_all[,c9], batch=sce_all[,c9]$run, subset.row = vargenes9, get.variance = TRUE)

set.seed(200)
sce_all_9_clust <- clusterCells(sce_all_9,
                                        use.dimred="corrected",
                                        BLUSPARAM=SNNGraphParam(k=20, 
                                                                type="jaccard",
                                                                cluster.fun="fast_greedy"))

table(sce_all_9_clust)
sce_all_9$cluster <- sce_all_9_clust
clust9 <- rep(0, times=ncol(sce_all_corrected))
clust9[c9] <- sce_all_9_clust

genes <- rownames(sce_all)[rowData(sce_all)$Symbol %in% c('CD8A', 'CD3E', 'KLRB1', 'CD8B')]
# note CD4 was not in the list of HVGs
assays(sce_all_9)$logcounts <- assays(sce_all)$logcounts[rownames(sce_all_9),c9]
plotExpression(sce_all_9, features=genes, x='cluster', exprs_values='logcounts')


```

This yields 3 subclusters, one which appears to be mainly CD3E/CD8A+, one which appears to have high KLRB1, and the other which is intermediate and likely marked by other genes. We'll go with this level of unsupervised subclustering so that we can extract the T cells and will have to think about the identities of the other subclusters later as well. 

```{r}

sce_all_9_clust <- as.character(sce_all_9_clust)
sce_all_9_clust[sce_all_9_clust == '1'] <- '9a'
sce_all_9_clust[sce_all_9_clust == '2'] <- '9b'
sce_all_9_clust[sce_all_9_clust == '3'] <- '9c'
cs <- as.character(sce_all_corrected$cluster)
cs[c9] <- sce_all_9_clust
sce_all_corrected$cluster <- cs

```

Finally, we need to re-run our marker scoring because we have added some clusters. For this, we will only write to file the scores for the clusters we have added.

```{r}
marker.info.perbatch <- scoreMarkers(sce_all_corrected, 
                                     groups=sce_all_corrected$cluster,
                                     block=sce_all_corrected$run,
                                     assay.type='logcounts',
                                     full.stats=TRUE)


for(i in 1:length(marker.info.perbatch)){
  if(names(marker.info.perbatch)[i] %in% c('18a', '18b', '20a', '20b', '9a', '9b', '9c')){
    ord <- marker.info.perbatch[[i]][order(marker.info.perbatch[[i]]$mean.logFC.cohen, decreasing=TRUE),]
    ord$symbol <- rowData(sce_all)[rownames(ord), 'Symbol']
    write.table(ord, 
                paste0('Results/cluster', names(marker.info.perbatch)[i], 'markers.txt'), 
                quote=F)
  }
}

```

And we make some plots now that we have all of the clusters.


```{r}
dat <- as.data.frame(table(sce_all_corrected$cluster, sce_all_corrected$Sample))
colnames(dat) <- c('cluster', 'sample', 'n_cells')

ggplot(dat, aes(fill=sample, y=n_cells, x=cluster)) +
  geom_bar(position='stack', stat='identity') + 
  theme_classic()

pdf('Results/cluster_occupancy.pdf')
ggplot(dat, aes(fill=sample, y=n_cells, x=cluster)) +
  geom_bar(position='stack', stat='identity') +
  theme_classic()
dev.off()

plotUMAP(sce_all_corrected, colour_by='cluster', point_size=0.3) +
  theme_classic()

pdf('Results/UMAP_clusters.pdf')
plotUMAP(sce_all_corrected, colour_by='cluster', point_size=0.3) + 
  theme_classic()
dev.off()

```

# Wrapping up

Write a data file to use in a shiny app for annotation.

```{r}
# for easier visualization in app, pre-convert counts metrics to log scale
sce_all_corrected$log2_sum_counts <- log2(sce_all_corrected$sum_counts)
sce_all_corrected$log2_genes_detected <- log2(sce_all_corrected$genes_detected)
# saveRDS(sce_all_corrected, file='shiny_app/data.rds')
```

Save everything

```{r}
sessionInfo()
save(list=ls(), file='ongoing_full_clustered.RData')
```

