---
title: "co-expression"
author: "Andrew_Muir"
date: "8/22/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

Following on from 'differential all'(Cross-ref is skipped). This script delineates patterns of gene co-expression between our experimental conditions (differential co-expression) with the end goal of revealing putative causal networks. 

**Excess of 240GB of RAM required to run**

# Load packages and data

Packages
```{r load-packages}
library(SingleCellExperiment)
library(scater)
library(scran)
library(dcanr)
library(igraph)
library(EBcoexpress)
#library(GeneNet)#ggm dropped for slowness and poor performance
#library(COSINE)#Not available on the older R version used by compute cluster

set.seed(123)
```

Data
```{r load-data}
load('Data/sce_all_filt.RData')
```

Trim genes to be investigated. Either DE or HVGs. 
```{r gene-lists}
dim(sce_all_filt)
unique(sce_all_filt$condition)

# Remove forbidden rowData names (for generation of CompressedGRangesList)
names(rowData(sce_all_filt))[which(names(rowData(sce_all_filt))=="start")]="gene_start"
names(rowData(sce_all_filt))[which(names(rowData(sce_all_filt))=="end")]="gene_end"

# Swap ENSEMBL IDs for HGNC where possible for easy visualization
library(foreach)
library(doParallel)
#Assuming more than 2 cores, leaving 2 cores free
numCores <- detectCores(logical=FALSE) -2
registerDoParallel(numCores) 

BAL_genes_mix <- vector("list", length(rownames(sce_all_filt)))

BAL_genes_mix <- foreach (n=1:length(rownames(sce_all_filt))) %dopar%{
  if(rowData(sce_all_filt)$hgnc_symbol[n]==""){
    BAL_genes_mix[n] <- rownames(sce_all_filt)[n]
  } else {
    BAL_genes_mix[n] <- rowData(sce_all_filt)$hgnc_symbol[n]
  }}
stopImplicitCluster()

BAL_genes_mix <- as.character(BAL_genes_mix)
rownames(sce_all_filt) <- BAL_genes_mix

```

pipelines for quick iteration and comparison of methods
```{r DC-pipelines}
#custom pipeline function
createDCnet <- function(emat, condition, dc.method, cor.method='spearman', ...) {
  #compute scores
  score = dcScore(logcounts(emat), condition, dc.method, cor.method=cor.method, ebcoexpress.useBWMC=FALSE,  ...)
  #perform statistical test
  pvals = dcTest(score, as.matrix(logcounts(emat)), condition, ...)
  #adjust tests for multiple testing
  adjp = dcAdjust(pvals, ...)
  #threshold and generate network
  dcnet = dcNetwork(score, adjp, ...)

  return(dcnet)
}
dcMethods()
# Function to return ranked differentially expressed genes by cluster
DEs_byCluster <- findMarkers(sce_all_filt, groups=colData(sce_all_filt)$cluster)

get_cluster_genes <- function(cutoff, genelist=TRUE){
  DEs_byCluster_shortlist <- NULL
  DEs_byCluster_clusterlist <- NULL
  for (n in 1:length(DEs_byCluster)){
    top_genes <- rownames(DEs_byCluster[[n]])[1:cutoff]
    DEs_byCluster_shortlist <- c(DEs_byCluster_shortlist, top_genes)
    cluster_list <- rep(names(DEs_byCluster)[n], length(top_genes))
    DEs_byCluster_clusterlist <- c(DEs_byCluster_clusterlist, cluster_list)
  }
  DEs_byCluster_df <- data.frame(matrix(NA, nrow = cutoff*length(DEs_byCluster), ncol = 2))
  colnames(DEs_byCluster_df) <- c('cluster', 'top_genes')
  DEs_byCluster_df$cluster <- DEs_byCluster_clusterlist
  DEs_byCluster_df$top_genes <- DEs_byCluster_shortlist
  DEs_byCluster_shortlist <- unique(DEs_byCluster_shortlist)
  if (genelist==FALSE){
    return(DEs_byCluster_df)
  } else {
    return(DEs_byCluster_shortlist)
  }
}
# Function to return ranked differentially expressed genes by condition
DEs_byCondition <- findMarkers(sce_all_filt, groups=colData(sce_all_filt)$condition)

get_condition_genes <- function(cutoff, genelist=TRUE){
  DEs_byCondition_shortlist <- NULL
  DEs_byCondition_Conditionlist <- NULL
  for (n in 1:length(DEs_byCondition)){
    top_genes <- rownames(DEs_byCondition[[n]])[1:cutoff]
    DEs_byCondition_shortlist <- c(DEs_byCondition_shortlist, top_genes)
    Condition_list <- rep(names(DEs_byCondition)[n], length(top_genes))
    DEs_byCondition_Conditionlist <- c(DEs_byCondition_Conditionlist, Condition_list)
  }
  DEs_byCondition_df <- data.frame(matrix(NA, nrow = cutoff*length(DEs_byCondition), ncol = 2))
  colnames(DEs_byCondition_df) <- c('Condition', 'top_genes')
  DEs_byCondition_df$Condition <- DEs_byCondition_Conditionlist
  DEs_byCondition_df$top_genes <- DEs_byCondition_shortlist
  DEs_byCondition_shortlist <- unique(DEs_byCondition_shortlist)
  if (genelist==FALSE){
    return(DEs_byCondition_df)
  } else {
    return(DEs_byCondition_shortlist)
  }
}

```
Edges are coloured based on the score (negative to positive scores are represented using the purple to green gradient of colours).

all genes and cells to get broad overview of networks between conditions **Heavy compute**
```{r broad-view}
genes  <- get_condition_genes(1000)

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("pH1N1", "PBS")]
dataSCE <- dataSCE[rownames(dataSCE) %in% genes, ]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, PBS vs infected")

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("Ad-HA/NP", "PBS")]
dataSCE <- dataSCE[rownames(dataSCE) %in% genes, ]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, PBS vs vaccinated")

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("Ad-HA/NP+Ad-IL1b", "PBS")]
dataSCE <- dataSCE[rownames(dataSCE) %in% genes, ]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, PBS vs vaccinated+IL-1b")

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("Ad-HA/NP+Ad-IL1b", "Ad-HA/NP")]
dataSCE <- dataSCE[rownames(dataSCE) %in% genes, ]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, vaccinated vs vaccinated+IL-1b")

```

all genes, but only in Tregs
```{r Tregs1}
Tregs_dataSCE <- sce_all_filt[, sce_all_filt$cluster %in% c("10")]
# Remove zero genes
Tregs_dataSCE = Tregs_dataSCE[rowSums(counts(Tregs_dataSCE)) != 0, ]

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("pH1N1", "PBS")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, PBS vs infected")

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("Ad-HA/NP", "PBS")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, PBS vs vaccinated")

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("Ad-HA/NP+Ad-IL1b", "PBS")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, PBS vs vaccinated+IL-1b")

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("Ad-HA/NP+Ad-IL1b", "Ad-HA/NP")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, vaccinated vs vaccinated+IL-1b")
```

all genes, all cells as above, but only visualising 1st neighbors of IFI6
```{r IFI6}
dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("pH1N1", "PBS")]
genes  <- get_condition_genes(1000)
dataSCE <- dataSCE[rownames(dataSCE) %in% genes, ]

test1 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore")
sub_g = make_ego_graph(test1, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test1, vertex.label = '', main="All genes, all cells, zscore")
plot(sub_g, vertex.label = '', main="IFI6 1st neighbors")

```

