---
title: "co-expression"
author: "Andrew_Muir"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

Following on from 'differential all'. This script delineates patterns of gene co-expression between our experimental conditions (differential co-expression) with the end goal of revealing putative causal networks. 

**256GB RAM recommended**

# Load packages and data

Packages
```{r load-packages}
library(SingleCellExperiment)
library(scater)
library(scran)
library(batchelor)
library(dcanr)
library(igraph)
library(EBcoexpress)

set.seed(123)
```

Data
```{r load-data}
load('Data/sce_all_filt.RData')
```

Processing; cell subset, normalise, gene subset, batch correction.
```{r preprocessing}
dim(sce_all_filt)
unique(sce_all_filt$condition)

# Remove forbidden rowData names (for generation of CompressedGRangesList)
names(rowData(sce_all_filt))[which(names(rowData(sce_all_filt))=="start")]="gene_start"
names(rowData(sce_all_filt))[which(names(rowData(sce_all_filt))=="end")]="gene_end"

# Restrict to selected cell types
sce_all_filt <- sce_all_filt[, sce_all_filt$cluster %in% c("1", "4", "8" , "12", "16", "7"," 20a", "10", "18b", "6", "20b", "18a", "9a", "9b", "9c")]

#Redo normalisation as logcounts now that clusters have been removed
sce_all_filt <- computePooledFactors(sce_all_filt, clusters=sce_all_filt$cluster, assay.type="counts")
sce_all_filt <- logNormCounts(sce_all_filt)

# Get top HVGs, batch corrected for sex, subset dataset
varmod <- modelGeneVar(sce_all_filt, block=sce_all_filt$sex)
top_HVGs <- getTopHVGs(varmod, n=1000)
sce_all_filt <- sce_all_filt[rownames(sce_all_filt) %in% top_HVGs, ]

# Batch Correction
mnnCorrected <- mnnCorrect(sce_all_filt, batch=sce_all_filt$run, k=20)
assay(sce_all_filt, i="corrected") <- assay(mnnCorrected, i="corrected")

# Swap ENSEMBL IDs for HGNC where possible for easy visualization
rownames(sce_all_filt) <- rowData(sce_all_filt)$Symbol

```

save post-processing chunk for convenience
```{r pre-processing-save}
save(list=ls(), file='Data/diff-co_prep.RData')
```

pipelines for quick iteration and comparison of methods
```{r DC-pipelines}
#custom pipeline function
createDCnet <- function(emat, condition, dc.method, cor.method='spearman', ...) {
  #compute scores
  score = dcScore(assay(emat, i="corrected"), condition, dc.method, cor.method=cor.method, ebcoexpress.useBWMC=FALSE,  ...)
  #perform statistical test
  pvals = dcTest(score, as.matrix(assay(emat, i="corrected")), condition, ...)
  #adjust tests for multiple testing
  adjp = dcAdjust(pvals, ...)
  #threshold and generate network
  dcnet = dcNetwork(score, adjp, ...)

  return(dcnet)
}
dcMethods()
# Function to return ranked differentially expressed genes by cluster
DEs_byCluster <- findMarkers(sce_all_filt, groups=colData(sce_all_filt)$cluster)

get_cluster_genes <- function(cutoff, genelist=TRUE){
  DEs_byCluster_shortlist <- NULL
  DEs_byCluster_clusterlist <- NULL
  for (n in 1:length(DEs_byCluster)){
    top_genes <- rownames(DEs_byCluster[[n]])[1:cutoff]
    DEs_byCluster_shortlist <- c(DEs_byCluster_shortlist, top_genes)
    cluster_list <- rep(names(DEs_byCluster)[n], length(top_genes))
    DEs_byCluster_clusterlist <- c(DEs_byCluster_clusterlist, cluster_list)
  }
  DEs_byCluster_df <- data.frame(matrix(NA, nrow = cutoff*length(DEs_byCluster), ncol = 2))
  colnames(DEs_byCluster_df) <- c('cluster', 'top_genes')
  DEs_byCluster_df$cluster <- DEs_byCluster_clusterlist
  DEs_byCluster_df$top_genes <- DEs_byCluster_shortlist
  DEs_byCluster_shortlist <- unique(DEs_byCluster_shortlist)
  if (genelist==FALSE){
    return(DEs_byCluster_df)
  } else {
    return(DEs_byCluster_shortlist)
  }
}
# Function to return ranked differentially expressed genes by condition
DEs_byCondition <- findMarkers(sce_all_filt, groups=colData(sce_all_filt)$condition)

get_condition_genes <- function(cutoff, genelist=TRUE){
  DEs_byCondition_shortlist <- NULL
  DEs_byCondition_Conditionlist <- NULL
  for (n in 1:length(DEs_byCondition)){
    top_genes <- rownames(DEs_byCondition[[n]])[1:cutoff]
    DEs_byCondition_shortlist <- c(DEs_byCondition_shortlist, top_genes)
    Condition_list <- rep(names(DEs_byCondition)[n], length(top_genes))
    DEs_byCondition_Conditionlist <- c(DEs_byCondition_Conditionlist, Condition_list)
  }
  DEs_byCondition_df <- data.frame(matrix(NA, nrow = cutoff*length(DEs_byCondition), ncol = 2))
  colnames(DEs_byCondition_df) <- c('Condition', 'top_genes')
  DEs_byCondition_df$Condition <- DEs_byCondition_Conditionlist
  DEs_byCondition_df$top_genes <- DEs_byCondition_shortlist
  DEs_byCondition_shortlist <- unique(DEs_byCondition_shortlist)
  if (genelist==FALSE){
    return(DEs_byCondition_df)
  } else {
    return(DEs_byCondition_shortlist)
  }
}

```
Edges are coloured based on the score (negative to positive scores are represented using the purple to green gradient of colours).

broad overview of networks between conditions
```{r broad-view}
dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("pH1N1", "PBS")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, PBS vs infected")

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("Ad-HA/NP", "PBS")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, PBS vs vaccinated")

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("Ad-HA/NP+Ad-IL1b", "PBS")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, PBS vs vaccinated+IL-1b")

dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("Ad-HA/NP+Ad-IL1b", "Ad-HA/NP")]
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, all cells, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, all cells, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, all cells, vaccinated vs vaccinated+IL-1b")

```

Tregs
```{r Tregs1}
Tregs_dataSCE <- sce_all_filt[, sce_all_filt$cluster %in% c("10")]
# Remove zero genes
Tregs_dataSCE = Tregs_dataSCE[rowSums(counts(Tregs_dataSCE)) != 0, ]

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("pH1N1", "PBS")]
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, PBS vs infected")
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, PBS vs infected")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, PBS vs infected")

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("Ad-HA/NP", "PBS")]
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, PBS vs vaccinated")
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, PBS vs vaccinated")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, PBS vs vaccinated")

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("Ad-HA/NP+Ad-IL1b", "PBS")]
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, PBS vs vaccinated+IL-1b")
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, PBS vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, PBS vs vaccinated+IL-1b")

dataSCE <- Tregs_dataSCE[, Tregs_dataSCE$condition %in% c("Ad-HA/NP+Ad-IL1b", "Ad-HA/NP")]
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy"), vertex.label = '', main="entropy, all genes, Tregs, vaccinated vs vaccinated+IL-1b")
#plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic"), vertex.label = '', main="magic, all genes, Tregs, vaccinated vs vaccinated+IL-1b")
plot(createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore"), vertex.label = '', main="zscore, all genes, Tregs, vaccinated vs vaccinated+IL-1b")
```

visualising 1st neighbors of IFI6
```{r IFI6}
dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("pH1N1", "PBS")]

test1 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore")
sub_g1 = make_ego_graph(test1, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test1, vertex.label = '', main="All genes, all cells, zscore")
plot(sub_g1, vertex.label = '', main="IFI6 1st neighbors, zscore")

test2 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="entropy")
sub_g2 = make_ego_graph(test1, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test1, vertex.label = '', main="All genes, all cells, entropy")
plot(sub_g2, vertex.label = '', main="IFI6 1st neighbors, entropy")

test3 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="magic")
sub_g3 = make_ego_graph(test1, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test1, vertex.label = '', main="All genes, all cells, magic")
plot(sub_g3, vertex.label = '', main="IFI6 1st neighbors, magic")

#Note all methods appear to return same gene neighbors for IFI6

IFI6_neighbors <- names(sub_g1[1])

IFI6_neighbors <- names(sub_g1["IFI6"])
length(rownames(sce_all_filt))
length(IFI6_neighbors)

```

IFI6 1st neighbors in macrophage, T cell and B cell subsets
```{r IFI6-subsets}
sub_dataSCE <- sce_all_filt[, sce_all_filt$condition %in% c("pH1N1", "PBS")]

dataSCE <- sub_dataSCE[, sub_dataSCE$cluster %in% c("1", "4", "8")]
test1 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore")
sub_g1 = make_ego_graph(test1, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test1, vertex.label = '', main="All genes, macrophages, zscore")
plot(sub_g1, vertex.label = '', main="IFI6 1st neighbors, macrophages,  zscore")

dataSCE <- sub_dataSCE[, sub_dataSCE$cluster %in% c("12", "16")]
test2 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore")
sub_g2 = make_ego_graph(test2, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test2, vertex.label = '', main="All genes, B Cells, zscore")
plot(sub_g2, vertex.label = '', main="IFI6 1st neighbors, B Cells,  zscore")

dataSCE <- sub_dataSCE[, sub_dataSCE$cluster %in% c("6", "7")]
test3 <- createDCnet(emat=dataSCE, condition=dataSCE$condition, dc.method="zscore")
sub_g3 = make_ego_graph(test3, order=1, mode="in", node="IFI6", mindist=0)[[1]]
plot(test3, vertex.label = '', main="All genes, T Cells, zscore")
plot(sub_g3, vertex.label = '', main="IFI6 1st neighbors, T Cells,  zscore")

length(names(sub_g1["IFI6"]))
length(names(sub_g2["IFI6"]))
length(names(sub_g3["IFI6"]))

sum(names(sub_g1["IFI6"]) %in% names(sub_g2["IFI6"]))/length(names(sub_g1["IFI6"]))
sum(names(sub_g1["IFI6"]) %in% names(sub_g3["IFI6"]))/length(names(sub_g1["IFI6"]))
sum(names(sub_g2["IFI6"]) %in% names(sub_g3["IFI6"]))/length(names(sub_g2["IFI6"]))
```

# Wrapping up

Save everything

```{r}
sessionInfo()
save(list=ls(), file='workspace_diff-co.RData') 
```