---
title: "DifferentialAnalysesPigBAL"
author: "Arianne Richard"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

This script picks up where "clustering all.Rmd" left off. 

This script is for running differential analyses between conditions. Specifically, we will be looking for differential abundance of cells within each cluster, as well as differential expression of cells within each cluster, according to treatment condition. There are 4 conditions: PBS, Ad-HA/NP, Ad-HA/NP+Ad-IL1b, pH1N1. There are male and female mice for each group, but sex is almost completely confounded with sets of runs, so we are unable to account for both of these in the same model. We will address this problem in both differential abundance and differential expression analyses. 

# Loading the data

```{r}
library(SingleCellExperiment)
library(DropletUtils)
library(scran)
library(scater)
library(uwot)
library(biomaRt)
library(scds)
library(batchelor)
library(bluster)

load('ongoing_full_clustered.RData')

## size of saved data got a bit out of hand
rm(all, sce_list)

```


# Differential abundance analysis

This analysis will be run using a negative binomial test. Because clusters were computed on the dataset after merging and correcting for differences between runs, we will not include run as a covariate here. Instead, we will account for sex while testing differential abundance between conditions. 

We also wish to account for the sorting strategy that was used by only comparing clusters that would have fallen into the same sorting gate (a bit of a glorified FACS analysis really). For this reason, we will split the data into the sorted populations of macrophages/monocytes/DCs, CD4+ Ts, CD8b+ Ts, and B cells before looking for differential abundance within these groups. So here we consider all of our clusters and split them according to major cell types (and whether we need to exclude them for other reasons). We'll first have a look at some key marker genes as a sanity check before running.


```{r}
## first look at marker genes for sorting

sce_all$cluster <- sce_all_corrected$cluster

plotExpression(sce_all_corrected, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD4'], x='cluster', exprs_values='logcounts')
plotExpression(sce_all, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD4'], x='cluster')

plotExpression(sce_all_corrected, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD8B'], x='cluster', exprs_values='logcounts')
plotExpression(sce_all, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD8B'], x='cluster')

plotExpression(sce_all_corrected, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD163'], x='cluster', exprs_values='logcounts')
plotExpression(sce_all, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD163'], x='cluster')

plotExpression(sce_all_corrected, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD14'], x='cluster', exprs_values='logcounts')
plotExpression(sce_all, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD14'], x='cluster')

plotExpression(sce_all_corrected, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD3D'], x='cluster', exprs_values='logcounts')
plotExpression(sce_all, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD3D'], x='cluster')

plotExpression(sce_all_corrected, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD3E'], x='cluster', exprs_values='logcounts')
plotExpression(sce_all, features=rownames(sce_all)[rowData(sce_all)$Symbol == 'CD3E'], x='cluster')

```

From this, we determine that the following clusters fall into each category. Of note, our differential abundance analyses will take place within categories that are smaller or the same size as those used for sorting. In this way, the sorting and binning won't play a role in our results.

CD4: 7, 10, 18b, 20a

CD8: 6, 9a, 18a, 20b

Macrophage/Monocyte/DC: 1, 4, 8, 13, 15

B: 2, 3, 12, 16

This leaves clusters 9b, 9c, 14, 19 and 21 to be removed without consideration, as well as clusters 5, 11 and 17, which were determined to be either doublet or poor quality. 

After running each of the sorted groups separately, we'll also run a global differential abundance analysis that ignores the sorting strategy, just out of interest. In theory, if the same proportions of each sorted population were indeed pooled for sequencing, the results should be very similar and will give us a chance to look at the clusters that we were not easily able to assign to a sorted class.

```{r}

CD4s <- sce_all_corrected[,sce_all_corrected$cluster %in% c('7', '10', '18b', '20a')]

CD8s <- sce_all_corrected[,sce_all_corrected$cluster %in% c('6', '9a', '18a', '20b')]

MMDs <- sce_all_corrected[,sce_all_corrected$cluster %in% c('1', '4', '8', '13', '15')]

Bs <- sce_all_corrected[,sce_all_corrected$cluster %in% c('2', '3', '12', '16')]

all <- sce_all_corrected[,sce_all_corrected$cluster %in% c('1', '2', '3', '4', '6', '7', '8', '9a', '9b', '9c', '10', '12', '13', '14', '15', '16', '18a', '18b', '19', '20a', '20b', '21')]

subsets <- list(CD4s=CD4s, CD8s=CD8s, MMDs=MMDs, Bs=Bs, all=all)

```

And we run our DA tests. We'll consider a change in abundance with any treatment versus PBS in an ANOVA-like analysis (strictly speaking, I believe this is analysis-of-deviance-like).

```{r}
library(edgeR)

DA <- lapply(subsets, function(x){
  abundances <- table(x$cluster, x$Sample)
  abundances <- unclass(abundances)
  anno <- annot[colnames(abundances),]
  y.ab <- DGEList(abundances, samples=anno)
  y.ab$samples$Sex <- factor(y.ab$samples$Sex)
  y.ab$samples$Condition <- factor(y.ab$samples$Condition, levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'))
  design <- model.matrix(~Sex + Condition, y.ab$samples)
  y.ab <- estimateDisp(y.ab, design, trend="none")
  summary(y.ab$common.dispersion)
  plotBCV(y.ab, cex=1)
  fit.ab <- glmQLFit(y.ab, design, robust=TRUE, abundance.trend=FALSE)
  summary(fit.ab$var.prior)
  summary(fit.ab$df.prior)
  plotQLDisp(fit.ab, cex=1)
  res <- glmQLFTest(fit.ab, coef=3:5)
  restab <- res$table
  restab$FDR <- p.adjust(restab$PValue, method='fdr')
  return(restab)
})

DA
```

Interestingly, the only strong effect here is on the proportion of Treg in with vaccine + IL1b. Let's plot it to have a look.

```{r}
## proportion of total cells
props <- prop.table(table(sce_all_corrected$cluster, sce_all_corrected$Sample), margin=2)['10',]
proptab <- annot[,c('sample', 'Condition', 'Sex')]
proptab$proportions <- props[rownames(proptab)]
proptab$Condition <- factor(proptab$Condition, levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'))

ggplot(data=proptab, aes(x=Condition, y=proportions)) +
  geom_boxplot() +
  geom_point(aes(col=Sex), position='jitter')

## and print it

pdf('Results/Treg_prop_total.pdf', height=10, width=10)
ggplot(data=proptab, aes(x=Condition, y=proportions)) +
  geom_boxplot(na.rm=TRUE) +
  geom_point(aes(col=Sex), position='jitter') +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme_classic()
dev.off()


## proportion of CD4s
props <- prop.table(table(CD4s$cluster, CD4s$Sample), margin=2)['10',]
proptab <- annot[,c('sample', 'Condition', 'Sex')]
proptab$proportions <- props[rownames(proptab)]
proptab$Condition <- factor(proptab$Condition, levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'))

ggplot(data=proptab, aes(x=Condition, y=proportions)) +
  geom_boxplot() +
  geom_point(aes(col=Sex), position='jitter')

## and print it

pdf('Results/Treg_prop_CD4.pdf', height=10, width=10)
ggplot(data=proptab, aes(x=Condition, y=proportions)) +
  geom_boxplot(na.rm=TRUE) +
  geom_point(aes(col=Sex), position='jitter') +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme_classic()
dev.off()


## and another format for among CD4s
props <- prop.table(table(CD4s$cluster, CD4s$Sample), margin=2)
props <- data.frame(t(props))
colnames(props) <- c('sample', 'cluster', 'prop')
props$sample <- factor(props$sample, levels=c('sample1', 'sample8', 'sample11', 'sample2', 'sample12', 'sample5', 'sample6' ,'sample9', 'sample4', 'sample7', 'sample10'))

ggplot(props, aes(fill=cluster, y=prop, x=sample)) +
  geom_bar(position='stack', stat='identity')

## and print it
pdf('Results/CD4_cluster_proportions.pdf', height=5, width=5)
ggplot(props, aes(fill=cluster, y=prop, x=sample)) +
  geom_bar(position='stack', stat='identity') +
  theme_classic()
dev.off()


```

# Differential expression analysis

Now we move on to differential expression analysis. Here our decision about correcting for run versus sex is more complicated as we will be using the raw values which do exhibit run-specific effects. Comparing results from sex- versus run-corrected data revealed some Y-chromosome genes were being called as DE when we only corrected for run. This is a problem as the balance of M/F differs by condition (either 1v2 or 2v1). To avoid this, we'll have to correct for sex and accept that there is orthogonal noise generated by each run that will simply reduce our power.

Based on our examinations in the clustering script, we exclude cluster 5 as low quality and clusters 17 and 11 as doublets. 

```{r}

sce_filt <- sce_all_corrected

sce_filt <- sce_filt[,!sce_filt$cluster %in% c('5', '17', '11')]

plotUMAP(sce_filt, colour_by='cluster')

```

Now we test for differential expression. This analysis will be run as a pseudobulk analysis, pooling counts from all of the cells within each cluster for each sample before treating like bulk RNAseq data. Differential expression will then be analysed for each cluster. For this first pass, we will use and ANOVA-like analysis (again, more like ANODEV I believe) for each cluster to ask whether any condition deviates from the others. 

```{r}

sce_all_filt <- sce_all[,colnames(sce_filt)]
sce_all_filt$cluster <- sce_filt$cluster

summed <- aggregateAcrossCells(sce_all_filt, 
    id=colData(sce_filt)[,c("cluster", "Sample")])

summed.filt <- summed[,summed$ncells >= 10]

pseudoBulkDGEmod <- function(clust, sce, coeffs, filtexprsmin=20, filtexprstotal=40){
  sc <- sce[,sce$cluster==clust]
  y <- DGEList(counts=counts(sc), samples=colData(sc))
  if(length(which(table(y$samples$condition) < 2)) > 0){
    return('insufficient cells per sample')}else{
      keep <- filterByExpr(y, group=sc$condition, min.count=filtexprsmin, min.total.count=filtexprstotal)
      y <- y[keep,]
      y <- calcNormFactors(y)
      y$samples$condition <- factor(y$samples$condition, levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'))
      par(mfrow=c(2,3))
      for (i in seq_len(ncol(y))) {
        plotMD(y, column=i, main=clust)
      }
      plotMDS(cpm(y, log=TRUE), 
              labels=y$samples$sex, 
              col=c('black', 'blue', 'turquoise', 'red')[as.numeric(factor(y$samples$condition, levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1')))], main=clust)
      design <- model.matrix(~factor(sex) + factor(condition, levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1')), y$samples)
      tc <- tryCatch(estimateDisp(y, design), error=function(e) e, warning=function(w) w)
      if(is(tc, 'error') | is(tc, 'warning')){return(paste(tc))}else{
        y <- estimateDisp(y, design)
        plotBCV(y, main=clust)
        fit <- glmQLFit(y, design, robust=TRUE)
        plotQLDisp(fit)
        res <- glmQLFTest(fit, coef=coeffs)
        return(res$table)
      }
    }
}

clusts <- as.character(levels(factor(summed.filt$cluster)))
de.results.any <- vector('list', length(clusts))
names(de.results.any) <- clusts

for(i in 1:length(clusts)){
  de.results.any[[i]] <- pseudoBulkDGEmod(clusts[i], summed.filt, coeffs=3:5)
}


cls <- lapply(de.results.any, class)
de.results.any <- de.results.any[-which(cls %in% 'character')]


for(i in 1:length(de.results.any)){
  de.results.any[[i]]$FDR <- p.adjust(de.results.any[[i]]$PValue, method='fdr')
  de.results.any[[i]]$symbol <- rowData(sce_all_filt)[rownames(de.results.any[[i]]), 'Symbol']
  colnames(de.results.any[[i]])[1:3] <- c('logFC_Ad.HA.NP', 'logFC_Ad.HA.NP.Ad.IL1b', 'logFC_pH1N1')
  write.table(de.results.any[[i]], 
              file=paste0('Results/DE_all_cluster_', names(de.results.any)[i], '.txt'), 
              sep='\t', quote=FALSE)
}
for(i in 1:length(de.results.any)){
  print(head(de.results.any[[i]][de.results.any[[i]]$FDR < 0.05,]))
}


```

There aren't big shifts by condition but there are a handful of genes in most clusters that significantly vary in one or more conditions. Also clusters 9b, 13, 19, 20, and 21 did not have enough cells in each sample/cluster split. 

We would now like to formally test each condition against PBS and generate a list of genes significant in each of these. 

```{r}

de.results.AdHANP <- vector('list', length(clusts))
names(de.results.AdHANP) <- clusts

for(i in 1:length(clusts)){
  de.results.AdHANP[[i]] <- pseudoBulkDGEmod(clusts[i], summed.filt, coeffs=3)
}

cls <- lapply(de.results.AdHANP, class)
de.results.AdHANP <- de.results.AdHANP[-which(cls %in% 'character')]

for(i in 1:length(de.results.AdHANP)){
  de.results.AdHANP[[i]]$FDR <- p.adjust(de.results.AdHANP[[i]]$PValue, method='fdr')
  de.results.AdHANP[[i]]$symbol <- rowData(sce_all_filt)[rownames(de.results.AdHANP[[i]]), 'Symbol']
  write.table(de.results.AdHANP[[i]], 
              file=paste0('Results/DE_AdHANP_cluster_', names(de.results.AdHANP)[i], '.txt'), 
              sep='\t', quote=FALSE)
}
for(i in 1:length(de.results.AdHANP)){
  print(head(de.results.AdHANP[[i]][de.results.AdHANP[[i]]$FDR < 0.05,]))
}




de.results.AdHANPAdIL1b <- vector('list', length(clusts))
names(de.results.AdHANPAdIL1b) <- clusts

for(i in 1:length(clusts)){
  de.results.AdHANPAdIL1b[[i]] <- pseudoBulkDGEmod(clusts[i], summed.filt, coeffs=4)
}

cls <- lapply(de.results.AdHANPAdIL1b, class)
de.results.AdHANPAdIL1b <- de.results.AdHANPAdIL1b[-which(cls %in% 'character')]

for(i in 1:length(de.results.AdHANPAdIL1b)){
  de.results.AdHANPAdIL1b[[i]]$FDR <- p.adjust(de.results.AdHANPAdIL1b[[i]]$PValue, method='fdr')
  de.results.AdHANPAdIL1b[[i]]$symbol <- rowData(sce_all_filt)[rownames(de.results.AdHANPAdIL1b[[i]]), 'Symbol']
  write.table(de.results.AdHANPAdIL1b[[i]], 
              file=paste0('Results/DE_AdHANPAdIL1b_cluster_', names(de.results.AdHANPAdIL1b)[i], '.txt'), 
              sep='\t', quote=FALSE)
}
for(i in 1:length(de.results.AdHANPAdIL1b)){
  print(head(de.results.AdHANPAdIL1b[[i]][de.results.AdHANPAdIL1b[[i]]$FDR < 0.05,]))
}




de.results.pH1N1 <- vector('list', length(clusts))
names(de.results.pH1N1) <- clusts

for(i in 1:length(clusts)){
  de.results.pH1N1[[i]] <- pseudoBulkDGEmod(clusts[i], summed.filt, coeffs=5)
}

cls <- lapply(de.results.pH1N1, class)
de.results.pH1N1 <- de.results.pH1N1[-which(cls %in% 'character')]

for(i in 1:length(de.results.pH1N1)){
  de.results.pH1N1[[i]]$FDR <- p.adjust(de.results.pH1N1[[i]]$PValue, method='fdr')
  de.results.pH1N1[[i]]$symbol <- rowData(sce_all_filt)[rownames(de.results.pH1N1[[i]]), 'Symbol']
  write.table(de.results.pH1N1[[i]], 
              file=paste0('Results/DE_pH1N1_cluster_', names(de.results.pH1N1)[i], '.txt'), 
              sep='\t', quote=FALSE)
}
for(i in 1:length(de.results.pH1N1)){
  print(head(de.results.pH1N1[[i]][de.results.pH1N1[[i]]$FDR < 0.05,]))
}


```

Now we make a few figures for sanity checking and presentation purposes.

Plot IFI6 in all clusters with PBS or pH1N1

```{r}
# example volcano
IL1b.CD8 <- de.results.AdHANPAdIL1b$`6`
cols <- rep('black', times=nrow(IL1b.CD8))
cols[IL1b.CD8$FDR < 0.05] <- 'red'
plot(IL1b.CD8$logFC, -log10(IL1b.CD8$PValue), col=cols)




IFI6 <- 'ENSSSCG00000034570'

plot_resid_gg <- function(y, clust, regress_out, split_by, split_by_levels, col_by, gene, ...){
  loggene <- cpm(y, log=TRUE)[gene,y$samples$cluster==clust]
  mod <- lm(loggene~y$samples[y$samples$cluster==clust, regress_out])
  res <- resid(mod)
  tab <- data.frame(res=res, split=y$samples[y$samples$cluster==clust,split_by],
                    col=y$samples[y$samples$cluster==clust,col_by])
  tab$split <- factor(tab$split, levels=split_by_levels)
  ggplot(data=tab, aes(x=split, y=res)) +
    geom_boxplot(na.rm=TRUE) +
    geom_point(aes(col=col), position='jitter') +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
    theme_classic()
}

y <- DGEList(counts=counts(summed.filt), samples=colData(summed.filt))

plot_resid_gg(y, clust='6', regress_out='sex', split_by='condition', split_by_levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'), col_by='run', gene=IFI6)
plot_resid_gg(y, clust='6', regress_out='run', split_by='condition', split_by_levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'), col_by='sex', gene=IFI6)
plot_resid_gg(y, clust='6', regress_out='run', split_by='condition', split_by_levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'), col_by='run', gene=IFI6)

## this is interesting as it has highlighted why we had such different values when including run instead of sex in the model - it looks like one of the pH1N1 samples that was in its own run (other sample was excluded) ended up getting flattened.

library(gridExtra)

pdf('Results/IFI6_pH1N1_egs.pdf', height=10, width=20)
p1 <- plot_resid_gg(y, clust='6', regress_out='sex', split_by='condition', split_by_levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'), col_by='sex', gene=IFI6, main='cluster 6')
p2 <- plot_resid_gg(y, clust='2', regress_out='sex', split_by='condition', split_by_levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'), col_by='sex', gene=IFI6, main='cluster 2')
p3 <- plot_resid_gg(y, clust='1', regress_out='sex', split_by='condition', split_by_levels=c('PBS', 'Ad-HA/NP', 'Ad-HA/NP+Ad-IL1b', 'pH1N1'), col_by='sex', gene=IFI6, main='cluster 1')
grid.arrange(p1, p2, p3, ncol=3)
dev.off()

```


# Cluster annotation

Based on manual annotation from the clustered data performed using the generated tables and our shiny app, we'll give the clusters more intuitive names that can be transposed into figures.

```{r}
# Manual annotation revealed that cluster 19 are epithelial cells, so we'll remove them from further analysis
sce_all_filt <- sce_all_filt[,!sce_all_filt$cluster %in% c('19')]

## add in cluster names as sce_all_filt$named_cluster
#Long-form annotated version including cell subtypes
for (i in unique(sce_all_filt$cluster)){
 if(any(i==c('7', '20a')))
    name <- 'CD4 T Cells' 
 if(any(i==c('18b')))
    name <- 'Mitotic CD4 T Cells'
 if(any(i==c('10')))
    name <- 'Tregs'
 if(any(i==c('6', '9a', '20b')))
    name <- 'CD8 T Cells' 
 if(any(i==c('18a')))
    name <- 'Mitotic CD8 T Cells'
 if(any(i==c('1', '4', '8')))
    name <- 'Macrophages' 
 if(any(i==c('15')))
    name <- 'Monocytes'
 if(any(i==c('13')))
    name <- 'cDCs'
 if(any(i==c('2', '3')))
    name <- 'Plasma Cells' 
 if(any(i==c('12', '16')))
    name <- 'B Cells' 
 if(any(i==c('14')))
    name <- 'T-cell-like' 
 if(any(i==c('9b', '9c')))
    name <- 'NK cells' 
 if(any(i==c('21')))
    name <- 'Unknown'
  sce_all_filt$named_cluster[sce_all_filt$cluster==i] <- name
}
# create 'main_clusters' column that ignores subclustering
main_clusters <- gsub("a", "", sce_all_filt$cluster)
main_clusters <- gsub("b", "", main_clusters)
main_clusters <- gsub("c", "", main_clusters)
sce_all_filt$main_clusters <- main_clusters

# Create label and colour slots for convenient annotation later
sce_all_filt$verbose_cluster <- paste0('BAL, ', sce_all_filt$named_cluster, ', ', sce_all_filt$cluster)
sce_all_filt$EXP_celltype <- paste0('BAL, ', sce_all_filt$named_cluster)

BAL_colour_list <- vector("list", length(sce_all_filt$EXP_celltype))
BAL_colour_list[grep('Macrophages', sce_all_filt$EXP_celltype)] <- '#E31A1C' #red
BAL_colour_list[grep('Monocytes', sce_all_filt$EXP_celltype)] <- '#FB9A99' #pink
BAL_colour_list[grep('cDCs', sce_all_filt$EXP_celltype)] <- '#FFFF99' #pale yellow
BAL_colour_list[grep('Plasma Cells', sce_all_filt$EXP_celltype)] <- '#1F78B4' #dark blue
BAL_colour_list[grep('B Cells', sce_all_filt$EXP_celltype)] <- '#A6CEE3' #pale blue
BAL_colour_list[grep('Tregs', sce_all_filt$EXP_celltype)] <- '#CAB2D6' #light purple
BAL_colour_list[grep('CD4', sce_all_filt$EXP_celltype)] <- '#B2DF8A' #light green
BAL_colour_list[grep('Mitotic CD4 T Cells', sce_all_filt$EXP_celltype)] <- '#FDBF6F' #pale orange
BAL_colour_list[grep('CD8', sce_all_filt$EXP_celltype)] <- '#33A02C' # green
BAL_colour_list[grep('Mitotic CD8 T Cells', sce_all_filt$EXP_celltype)] <- '#FF7F00' #orange
BAL_colour_list[grep('NK', sce_all_filt$EXP_celltype)] <- '#6A3D9A' #purple
BAL_colour_list[sapply(BAL_colour_list, is.null)] <- '#B15928' #brown
sce_all_filt$colour  <- BAL_colour_list

legend_colours <- BAL_colour_list
names(legend_colours) <- sce_all_filt$named_cluster
legend_colours <- legend_colours[!duplicated(names(legend_colours))]
legend_colours <- legend_colours[c("Macrophages", "Monocytes", "cDCs", "Plasma Cells", "B Cells", "CD4 T Cells", "Tregs", "Mitotic CD4 T Cells", "Mitotic CD8 T Cells", "CD8 T Cells", "NK cells", "T-cell-like", "Unknown")]

df_legend_colours <- data.frame(unlist(names(legend_colours)), unlist(legend_colours))
names(df_legend_colours) <- c("Cell_Type", "Colour")
df_legend_colours$dummy_data <- seq(1, length(df_legend_colours[,1]))

# Plot new named clusters
png('Results/UMAP_named_clusters.png', height=1080, width=2160)
p1 <- plotUMAP(sce_all_filt, colour_by="cluster", by_exprs_values='reconstructed', point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
p2 <- plotUMAP(sce_all_filt, colour_by="named_cluster", by_exprs_values='reconstructed', point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
grid.arrange(p1, p2, ncol=2)
dev.off()

png('Results/UMAP_named_and_numbered_clusters.png', height=2160, width=2160)
plotUMAP(sce_all_filt, colour_by="cluster", text_by="main_clusters", text_size=20, by_exprs_values='reconstructed', point_size=0.3) +
guides(colour = guide_legend(override.aes = list(size=30)))+ theme(text = element_text(size = 40), legend.title=element_blank(), legend.position=c(0.1,0.3))
dev.off()

library(gridExtra)

png(filename='Results/UMAP_named_and_numbered_clusters.png', height=1080, width=2160)
plot0 <- ggplot(data=df_legend_colours, aes(x=dummy_data)) +
  geom_point(aes(y=dummy_data, color=Cell_Type), size=10) +
  labs(x=NULL, y=NULL, color=NULL) +
  scale_color_manual(values = legend_colours) +
  guides(colour = guide_legend(ncol = 1)) + 
  theme(legend.position=c(0.5,0), legend.direction="vertical", aspect.ratio=0, text = element_text(size = 50),
        axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.text.y=element_blank(), axis.ticks.y=element_blank())
plot1 <- plotUMAP(sce_all_filt, point_size=0, text_by="main_clusters", text_size=20) +
  aes(colour = sce_all_filt$colour) +
  geom_point(shape='.', size=1, alpha=0.5) +
  geom_text(x=-10, y=8, size=30, label="B", color="black") +
  theme(aspect.ratio = 1, plot.margin = unit(c(0,0,0,0), "cm"), text = element_text(size = 40))
grid.arrange(plot1, plot0, ncol=2)
dev.off()

# Update shiny app to filtered data for easier visualization and identification
sce_all_filt$log2_sum_counts <- log2(sce_all_filt$sum_counts)
sce_all_filt$log2_genes_detected <- log2(sce_all_filt$genes_detected)
saveRDS(sce_all_filt, file='shiny_app/data.rds')

```

# Wrapping up

Save everything

```{r}
sessionInfo()
save(sce_all_filt, file='Data/sce_all_filt.RData')
save(list=ls(), file='ongoing_full_diff.RData') 
```



