---
title: "Cross_ref"
author: "Andrew_Muir"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

Continuing from 'differential all.Rmd'.

This script performs a familiarity analysis between the clusters and cell types identified so far in the PirbrightBAL analysis against the reference transcriptome published by Herrera-Uribe et al. 2021 (https://doi.org/10.3389%2Ffgene.2021.689406). 

This includes a mix of cell-type mapping, cluster mapping and cell to cell mapping. 

**Note: This script uses 50+ GB of memory, run on a compute cluster!**

# Load packages and data
pigBALsce is from 'differential all.Rmd'.
PBMC's are from:
https://data.nal.usda.gov/dataset/data-reference-transcriptomics-porcine-peripheral-immune-cells-created-through-bulk-and-single-cell-rna-sequencing/resource/7c096891-2807-4e4e-8f5f-0c02c6f434c4

Packages
```{r load-packages}
library(SingleCellExperiment)
library(Seurat)
library(SeuratDisk)
library(scmap)
library(scater)
library(scran)
library(batchelor)
```
Data
```{r load-data}
load('Data/sce_all_filt.RData')
pigBALsce <- sce_all_filt
rm(sce_all_filt)

pigPBMC <- LoadH5Seurat('Herrera-Uribe/PBMC7_AllCells.h5seurat')
```
OR load workspace from previous session. 
```{r load-workspace}
#load('Data/CellTypeSave.RData') #post pre-processing and cell type validation
#load('Data/ClusterSave.RData') #post scmap cluster mapping
#load('Data/mnnCorrectSave.RData') #Post mnnCorrection
```

# Basic comparison of datasets

```{r dataset-metrics}
#Check size and depth
dim(pigBALsce)
dim(pigPBMC)

#Check identified cell types
unique(pigBALsce$named_cluster)
unique(pigPBMC$celltypes)

#Check number of clusters
unique(pigBALsce$cluster)
unique(pigPBMC$seurat_clusters)

```
BAL dataset has 3000 more genes and twice as many cells as PBMC. 

Some good overlap in the important cell types, though the BAL has several cell subtypes that are combined in the PBMC dataset (e.g. mitotic T cells and plasma cells). Obviously this is partially due to the different tissues. We'll deal with this later. 

19 vs 36 clusters, though BAL has had doublets/low-quality clusters removed post clustering. Again deal with synchronization later.

# Check alignment and gene annotation
Both datasets are aligned to the same pig genome from ENSEMBL, albeit different release versions of ENSEMBL (V97 vs V107) and used different alignment functions. This shouldn't be an issue, though we'll do a quick check just in case. 

pigBAL
(cellranger-7.0.0) using default parameters and Sus scrofa genome (genome assembly 11.1, Ensembl release 107).
Pirbright pigs

pigPBMC
Sscrofa 11.1 (V97; http://ftp.ensembl.org/pub/release-97/fasta/sus_scrofa/dna/) and annotation GTF (v11.1.97; http://ftp.ensembl.org/pub/release-97/gtf/sus_scrofa/)
custom Perl scripts
Large White and Landrace pigs

Incidentally, both Landrace and Large White pigs were included in the nine breeds used to generate genome assembly 11.1, Pirbright pigs were not. 

```{r check-initial-shared-genes}
BAL_genes <- rowData(pigBALsce)$hgnc_symbol
PBMC_genes <- rownames(pigPBMC@assays$RNA@counts)
# % BAL genes in PBMC
sum(BAL_genes %in% PBMC_genes)/nrow(pigBALsce)
# % PBMC genes in BAL
sum(PBMC_genes %in% BAL_genes)/nrow(pigPBMC)

# Quick check to see if ENSEMBL IDs work better
sum(rownames(pigBALsce) %in% rownames(pigPBMC@assays$RNA@counts)/length(rownames(pigBALsce)))

```
56% and 68% shared genes, not great but not a disaster either. Lets see if there's a a trend in the genes not shared. 

```{r check-unshared-genes}
length(BAL_genes[(BAL_genes %in% PBMC_genes)])

head(BAL_genes[!(BAL_genes %in% PBMC_genes)], n=50)

head(PBMC_genes[!(PBMC_genes %in% BAL_genes)], n=50)

# Quick check to see if ENS IDs have been updated with new ENSEMBL releases
old_ENS <- grep('ENS', PBMC_genes[!(PBMC_genes %in% BAL_genes)], value=TRUE)
```
Cross checked using ENSEMBL ID history checker, IDs are all fine. They shouldn't have been altered by ENSEMBL releases, only new genome builds (both datasets have same genome). 

BAL gene annotations (for HGNC) include "" where there are no HGNC symbols. PBMC gene annotations are a mix of ENSEMBL IDs and HGNC symbols (single slot). This may be the cause of the problem. 

Next we'll try updating the HGNC symbols for both datasets, otherwise we'll need to try realignment.  

**Note:this chunk exports and imports the workload elsewhere so I could get a faster internet connection.**
```{r update-HGNC}
save(BAL_genes, file = 'Data/BAL_genes.RData')
save(PBMC_genes, file = 'Data/PBMC_genes.RData')
##Ran elsewhere
#new_symbols_BAL <- UpdateSymbolList(symbols = BAL_genes)
#save(new_symbols_BAL, file = 'Data/new_symbols_BAL.RData')
#new_symbols_PBMC <- UpdateSymbolList(symbols = PBMC_genes)
#save(new_symbols_PBMC, file = 'Data/new_symbols_PBMC.RData')
##Load results
load('Data/new_symbols_BAL.RData')
load('Data/new_symbols_PBMC.RData')
# Recheck genes shared
# % BAL genes in PBMC
sum(new_symbols_BAL %in% new_symbols_PBMC)/length(new_symbols_BAL)
# % PBMC genes in BAL
sum(new_symbols_PBMC %in% new_symbols_BAL)/length(new_symbols_PBMC)

#Some genes have been deleted, has this shifted placement?
tail(BAL_genes, 10)
tail(new_symbols_BAL, 10)
tail(PBMC_genes, 10)
tail(new_symbols_PBMC, 10)
```
136 and 326 genes replaced respectively. 57% and 69% genes now shared, a marginal (1%) improvement by updating the HGNC symbols. The updated symbol lists however are shorter than the originals, by a total of 6 genes, order of the genes appears the same. However I don't trust that the loss of six genes may have shifted the alignment with the original objects, hence I won't be using the updated HGNC symbols despite the 1% improvement. 

Another option is to combine the ENSEMBL IDs with HGNC symbols in the BAL dataset, to mimic the PBMCs. 
```{r mix-HGNC-with-ENSMEBL-for-BAL}
library(foreach)
library(doParallel)
#Assuming more than 2 cores, leaving 2 cores free
numCores <- detectCores(logical=FALSE) -2
registerDoParallel(numCores) 

BAL_genes_mix <- vector("list", length(rownames(pigBALsce)))

BAL_genes_mix <- foreach (n=1:length(rownames(pigBALsce))) %dopar%{
  if(rowData(pigBALsce)$hgnc_symbol[n]==""){
    BAL_genes_mix[n] <- rownames(pigBALsce)[n]
  } else {
    BAL_genes_mix[n] <- rowData(pigBALsce)$hgnc_symbol[n]
  }}
stopImplicitCluster()

BAL_genes_mix <- as.character(BAL_genes_mix)

# % BAL genes in PBMC
sum(BAL_genes_mix %in% PBMC_genes)/nrow(pigBALsce)
# % PBMC genes in BAL
sum(PBMC_genes %in% BAL_genes_mix)/nrow(pigPBMC)

# Now we can try updating the BAL genes again (run elsewhere again)
#updated_BAL_genes_mix <- UpdateSymbolList(symbols = BAL_genes_mix)
#save(updated_BAL_genes_mix, file='Data/updated_BAL_genes_mix.RData')
load('Data/updated_BAL_genes_mix.RData')

# % BAL genes in PBMC
sum(updated_BAL_genes_mix %in% new_symbols_PBMC)/length(updated_BAL_genes_mix)
# % PBMC genes in BAL
sum(new_symbols_PBMC %in% updated_BAL_genes_mix)/length(new_symbols_PBMC)
```
63% and 76% genes shared using the mixed BAL gene list, so this has improved by 8% (an extra 1200genes).

Updated HGNC symbols provides 64% and 77% shared genes (almost 1% improvement). However once again updating the HGNC symbols has deleted four genes, therefore I won't be using the updated gene list. 

# Quick check of HVGs
```{r HVG-check}
rownames(pigBALsce) <- BAL_genes_mix
top_HVGs_BAL <- getTopHVGs(pigBALsce, n=1000)
pigPBMC <- FindVariableFeatures(pigPBMC, nfeatures=1000)
top_HVGs_PBMC <- VariableFeatures(pigPBMC)

sum(top_HVGs_BAL %in% top_HVGs_PBMC)/length(top_HVGs_BAL)
sum(top_HVGs_PBMC %in% top_HVGs_BAL)/length(top_HVGs_PBMC)
```
Only 14% shared HGVs, not great but we'll see if it poses an issue later. 

# Standardisation for scmap
The first analysis we will be running is scmap, which requires that both datasets are sce, that data is stored in specific named slots and that some slots are cleared. 

```{r create-scmap-compatible-sce}
# Remove forbidden rowData names (for generation of CompressedGRangesList)
names(rowData(pigBALsce))[which(names(rowData(pigBALsce))=="start")]="gene_start"
names(rowData(pigBALsce))[which(names(rowData(pigBALsce))=="end")]="gene_end"

# Add slots to BAL
rowData(pigBALsce)$feature_symbol <- BAL_genes_mix
colData(pigBALsce)$cell_type1 <- pigBALsce$cluster

# Create PBMC sce from seurat object
pigPBMCsce <- as.SingleCellExperiment(pigPBMC)

# Add PBMC slots 
rowData(pigPBMCsce)$feature_symbol <- rownames(pigPBMC@assays$RNA@counts)
colData(pigPBMCsce)$cell_type1 <- as.character(pigPBMC$seurat_clusters)

# Remove genes with zero expression
pigBALsce = pigBALsce[rowSums(counts(pigBALsce)) != 0, ]
pigPBMCsce = pigPBMCsce[rowSums(counts(pigPBMCsce)) != 0, ]

# Restrict both datasets to shared genes only
length(rownames(pigBALsce))
pigBALsce <- pigBALsce[rowData(pigBALsce)$feature_symbol %in% rowData(pigPBMCsce)$feature_symbol]
length(rownames(pigBALsce))

length(rownames(pigPBMCsce))
pigPBMCsce <- pigPBMCsce[rowData(pigPBMCsce)$feature_symbol %in% rowData(pigBALsce)$feature_symbol]
length(rownames(pigPBMCsce))

# Create verbose named cluster slot for convenient annotation later
pigBALsce$verbose_cluster <- paste0('BAL, ', pigBALsce$named_cluster, ', ', pigBALsce$cluster)
pigPBMCsce$verbose_cluster <- paste0('PBMC, ', pigPBMCsce$celltypes, ', ', pigPBMCsce$seurat_clusters)

# Select gene features for scmap
rownames(pigPBMCsce) <- rowData(pigPBMCsce)$feature_symbol
rownames(pigBALsce) <- rowData(pigBALsce)$feature_symbol

BAL_feature_genes <- findMarkers(pigBALsce, groups=colData(pigBALsce)$cell_type1)
PBMC_feature_genes <- findMarkers(pigPBMCsce, groups=colData(pigPBMCsce)$cell_type1)

cutoff <- 50

BAL_gene_shortlist <- NULL
BAL_cluster_shortlist <- NULL
for (n in 1:length(BAL_feature_genes)){
  top_genes <- rownames(BAL_feature_genes[[n]])[1:cutoff]
  BAL_gene_shortlist <- c(BAL_gene_shortlist, top_genes)
  cluster_list <- rep(names(BAL_feature_genes)[n], length(top_genes))
  BAL_cluster_shortlist <- c(BAL_cluster_shortlist, cluster_list)
}
BAL_feature_genes_df <- data.frame(matrix(NA, nrow = cutoff*length(BAL_feature_genes), ncol = 2))
colnames(BAL_feature_genes_df) <- c('cluster', 'top_genes')
BAL_feature_genes_df$cluster <- BAL_cluster_shortlist
BAL_feature_genes_df$top_genes <- BAL_gene_shortlist
BAL_gene_shortlist <- unique(BAL_gene_shortlist)

PBMC_gene_shortlist <- NULL
PBMC_cluster_shortlist <- NULL
for (n in 1:length(PBMC_feature_genes)){
  top_genes <- rownames(PBMC_feature_genes[[n]])[1:cutoff]
  PBMC_gene_shortlist <- c(PBMC_gene_shortlist, top_genes)
  cluster_list <- rep(names(PBMC_feature_genes)[n], length(top_genes))
  PBMC_cluster_shortlist <- c(PBMC_cluster_shortlist, cluster_list)  
}
PBMC_feature_genes_df <- data.frame(matrix(NA, nrow = cutoff*length(PBMC_feature_genes), ncol = 2))
colnames(PBMC_feature_genes_df) <- c('cluster', 'top_genes')
PBMC_feature_genes_df$cluster <- PBMC_cluster_shortlist
PBMC_feature_genes_df$top_genes <- PBMC_gene_shortlist
PBMC_gene_shortlist <- unique(PBMC_gene_shortlist)
```

```{r visualise-new-sce-data}
#Sanity check, visualise the converted data, attempt to recreate published visuals
set.seed(123)
pigPBMCsce <- fixedPCA(pigPBMCsce, subset.row=NULL)
pigPBMCsce <- runUMAP(pigPBMCsce, dimred='PCA', n_dimred = 1:10, n_neighbors=20)
#UMAPs
plotUMAP(pigPBMCsce, colour_by="seurat_clusters", point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
plotUMAP(pigPBMCsce, colour_by="celltypes", point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
# Highlight T cells, check if correct annotation
plotUMAP(pigPBMCsce, colour_by="CD3E", point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
# Highlight monocytes, check if correct annotation
plotUMAP(pigPBMCsce, colour_by="CD14", point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
# Highlight B cells, check if correct annotation
plotUMAP(pigPBMCsce, colour_by="CD19", point_size=0.3) + guides(colour = guide_legend(override.aes = list(size=8)))+ theme(text = element_text(size = 20))
```
Plots aren't the cleanest in terms of cluster localisation (compared to published), however CD3E, CD14 and CD19 appear to localise to the correct cell types, so the conversion hasn't broken anything and we're okay to move on. 

```{r cell-type-save}
save(list=ls(), file='Data/CellTypeSave.RData') 
```

# scmap BAL cluster to PBMC cluster
select features on BAL
```{r select-features-BAL-clusters}
pigBALsce <- setFeatures(pigBALsce, BAL_gene_shortlist)
```
index on BAL
```{r BAL-cluster-index}
pigBALsce <- indexCluster(pigBALsce)
```
select features on PBMC
```{r select-features-PBMC-clusters}
pigPBMCsce <- setFeatures(pigPBMCsce, PBMC_gene_shortlist)
```
index on PBMC
```{r PBMC-cluster-index}
pigPBMCsce <- indexCluster(pigPBMCsce)
```
project data
```{r cluster-cluster-projection}
clusterBAL_to_clusterPBMC <- scmapCluster(
  projection = pigBALsce,
  index_list = list(pigPBMCsce = metadata(pigPBMCsce)$scmap_cluster_index),
  threshold = 0.7
)
```
create table
```{r cluster-cluster-tables}
cluster_clusterBAL_proj_table <- table(colData(pigBALsce)$cell_type1, clusterBAL_to_clusterPBMC$scmap_cluster_labs)
cluster_clusterBAL_proj_table
write.csv(cluster_clusterBAL_proj_table, file='Data/cluster-clusterBAL_projection.csv')

# Make percentage table
cluster_clusterBAL_percent_proj_table <- round(prop.table(cluster_clusterBAL_proj_table, margin=1),2)
cluster_clusterBAL_percent_proj_table
write.csv(cluster_clusterBAL_percent_proj_table, file='Data/cluster-clusterBAL_percent_projection.csv')
```
plot quick sankey
```{r cluster-cluster-sankey}
plot(getSankey(colData(pigBALsce)$cell_type1,  clusterBAL_to_clusterPBMC$scmap_cluster_labs[,1], colors=c('#FF0000', '#FF9900', '#99FF00', '#00FF99', '#0099FF', '#0000FF', '#9900FF',  '#FF0099')))
```

cluster to cluster works really well, better than any of the cell type implementations. May be better to only use cluster to cluster mapping and then annotate with defined cell types. 

# Annotate and truncate cluster to cluster mapping table, prep for networkD3 sankey

```{r map-table-visual}
map_table <- cluster_clusterBAL_proj_table
# re-order
map_table <- map_table[order(rownames(map_table), decreasing = FALSE), ]
map_table <- map_table[, order(colnames(map_table), decreasing = FALSE)]

#rename BAL
BAL_clusters <- unique(data.frame(pigBALsce$cluster, pigBALsce$named_cluster))
BAL_clusters <- BAL_clusters[match(rownames(map_table), BAL_clusters$pigBALsce.cluster),]

if(identical(rownames(map_table), BAL_clusters$pigBALsce.cluster) == FALSE){
  stop("BAL names not identical")
}
BAL_names <- paste0(BAL_clusters$pigBALsce.named_cluster, ', ', BAL_clusters$pigBALsce.cluster)
rownames(map_table) <- BAL_names

#rename PBMC
PBMC_clusters <- unique(data.frame(as.character(pigPBMCsce$seurat_clusters), as.character(pigPBMCsce$celltypes)))
PBMC_clusters <- PBMC_clusters[match(colnames(map_table), PBMC_clusters$as.character.pigPBMCsce.seurat_clusters),]
PBMC_clusters[is.na(PBMC_clusters)] <- 'unassigned'

if(identical(colnames(map_table), PBMC_clusters$as.character.pigPBMCsce.seurat_clusters) == FALSE){
  stop("PBMC names not identical")
}
PBMC_names <- paste0(PBMC_clusters$as.character.pigPBMCsce.celltypes, ', ', PBMC_clusters$as.character.pigPBMCsce.seurat_clusters)
colnames(map_table) <- PBMC_names

# re-order, now with new names
map_table <- map_table[order(rownames(map_table), decreasing = FALSE), ]
map_table <- map_table[, order(colnames(map_table), decreasing = FALSE)]
write.csv(map_table, file='Results/cluster_projection_counts.csv')

# as percentage
map_percent_table <- round(prop.table(map_table, margin=1),2)
write.csv(map_percent_table, file='Results/cluster_projection_percent.csv')

# Prep for sankey
map_dataframe <- data.frame(map_percent_table)
# Trim flow, remove values below 5%
map_dataframe <- map_dataframe[map_dataframe$Freq > 0.15, ]

```

# Create combined sankey
percentage based may be better for plotting cluster-cluster

```{r networkD3-sankey}
library(networkD3)
library(dplyr)
 
# connection data frame
links <- data.frame(
  source=map_dataframe$Var1, 
  target=map_dataframe$Var2, 
  value=map_dataframe$Freq
  )
 
# node data frame
nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)
# Group nodes by cell type
grouplist <- vector("list", length(nodes$name))

grouplist[grep('Plasma Cells', nodes$name)] <- 'a'
grouplist[grep('ASC', nodes$name)] <- 'a'
grouplist[grep('B Cells', nodes$name)] <- 'a'
grouplist[grep('B cells', nodes$name)] <- 'a'
grouplist[grep('CD4', nodes$name)] <- 'b'
grouplist[grep('Tregs', nodes$name)] <- 'b'
grouplist[grep('CD8', nodes$name)] <- 'c'
grouplist[grep('NK', nodes$name)] <- 'c'
grouplist[grep('Monocytes', nodes$name)] <- 'd'
grouplist[grep('Macrophages', nodes$name)] <- 'd'
grouplist[grep('DCs', nodes$name)] <- 'd'
grouplist[grep('CD2', nodes$name)] <- 'f'
grouplist[sapply(grouplist, is.null)] <- 'e'

nodes$group <- grouplist

# Group links by cell type
grouplist2 <- vector("list", length(links$source))

grouplist2[grep('Plasma Cells', links$source)] <- 'a'
grouplist2[grep('ASC', links$source)] <- 'a'
grouplist2[grep('B Cells', links$source)] <- 'a'
grouplist2[grep('B cells', links$source)] <- 'a'
grouplist2[grep('CD4', links$source)] <- 'b'
grouplist2[grep('Tregs', links$source)] <- 'b'
grouplist2[grep('CD8', links$source)] <- 'c'
grouplist2[grep('NK', links$source)] <- 'c'
grouplist2[grep('Monocytes', links$source)] <- 'd'
grouplist2[grep('Macrophages', links$source)] <- 'd'
grouplist2[grep('DCs', links$source)] <- 'd'
grouplist2[sapply(grouplist2, is.null)] <- 'e'

links$group <- grouplist2

# Give each group a colour
my_color <- 'd3.scaleOrdinal(d3.schemeCategory10) .domain(["a", "b", "c", "d", "e", "f"])'

# convert names to IDs
links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1
 
# Make the Network
sanPlot <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name",
              colourScale=my_color, LinkGroup="group", NodeGroup="group",
              fontFamily = "arial", fontSize = 8, nodeWidth = 8,
              sinksRight=FALSE)
sanPlot

htmltools::save_html(sanPlot, file = "Results/cluster_map_sankey.html")
```


# Identify key mapping genes
Use original gene counts (of selected features), subset by interesting comparisons (cell-types), sort, and plot as heatmaps.
```{r identify-map-genes}
# Re-order gene names for convenience
pigPBMCsce <- pigPBMCsce[match(rowData(pigBALsce)$feature_symbol, rowData(pigPBMCsce)$feature_symbol),]
identical(rownames(pigPBMCsce), rownames(pigBALsce))

# Normalise between datasets, useful for visualisation and later mnncorrect

library(scuttle)

# B cells
B_Cell_genes <- PBMC_feature_genes_df[grep(pattern=c("10|16"), x=PBMC_feature_genes_df$cluster), ]
B_Cell_genes <- unique(B_Cell_genes$top_genes)

BAL_B_cells <- pigBALsce[rownames(pigBALsce) %in% B_Cell_genes, ]
BAL_B_cells <- BAL_B_cells[, BAL_B_cells$cluster %in% c("12", "16", "19")]
PBMC_B_cells <- pigPBMCsce[rownames(pigPBMCsce) %in% B_Cell_genes, ]
PBMC_B_cells <- PBMC_B_cells[, PBMC_B_cells$seurat_clusters %in% c("10", "16")]
identical(rownames(PBMC_B_cells), rownames(BAL_B_cells))

BAL_B_cells <- as.matrix(logcounts(aggregateAcrossCells(BAL_B_cells, ids=BAL_B_cells$verbose_cluster, statistics='mean', use.assay.type='logcounts')))
PBMC_B_cells <- as.matrix(logcounts(aggregateAcrossCells(PBMC_B_cells, ids=PBMC_B_cells$verbose_cluster, statistics='mean', use.assay.type='logcounts')))

png(filename='Results/B_Cells_Cross_Heatmap.png', height=1080, width=2160)
heatmap(cbind(BAL_B_cells, PBMC_B_cells), margins=c(15,5), keep.dendro = FALSE)
dev.off()
```

Next we'll try mnnCorrect, which corrects for experimental differences to allow for joint or merged analysis. 

save
```{r cluster-save}
save(list=ls(), file='Data/ClusterSave.RData') 
```

# mnnCorrect
Genes in both datasets need to be in identical order. 
```{r standardise-for-mnnCorrect}
pigPBMCsce <- pigPBMCsce[match(rowData(pigBALsce)$feature_symbol, rowData(pigPBMCsce)$feature_symbol),]
head(rowData(pigBALsce)$feature_symbol, n=10)
head(rowData(pigPBMCsce)$feature_symbol, n=10)
tail(rownames(pigBALsce), n=10)
tail(rownames(pigPBMCsce), n=10)
identical(rownames(pigPBMCsce), rownames(pigBALsce))
identical(rowData(pigBALsce)$feature_symbol, rowData(pigPBMCsce)$feature_symbol)
```
Restricting genes fed to mnnCorrect to improve merging and speed up computation. 
```{r trim-genes-mnncorrect}
Union_feature_genes <- paste0(c(BAL_gene_shortlist, PBMC_gene_shortlist))
Union_feature_genes <- unique(Union_feature_genes)

BAL_Union_sce <- pigBALsce[rownames(pigBALsce) %in% Union_feature_genes, ]
PBMC_Union_sce <- pigPBMCsce[rownames(pigPBMCsce) %in% Union_feature_genes, ]
identical(rownames(BAL_Union_sce), rownames(PBMC_Union_sce))
```

Run mnnCorrect
**Warning: this step is compute heavy. Will take several days to complete.**
```{r run-mnnCorrect}
## Trying fast MNN to see how 'fast' it is. Not ideal since it reduces results to a PC space.
start_time <- Sys.time()
fastmnnCorrected <- fastMNN(logcounts(BAL_Union_sce), logcounts(PBMC_Union_sce), k=20)
end_time <- Sys.time()
fastmnn_time <- end_time-start_time
save(fastmnnCorrected, file='Data/fastmnnCorrected_feature_k20.RData')
#load('Data/fastmnnCorrected.RData')
start_time <- Sys.time()
mnnCorrected <- mnnCorrect(logcounts(BAL_Union_sce), logcounts(PBMC_Union_sce), k=20)
end_time <- Sys.time()
slowmnn_time <- end_time-start_time
slowmnn_time
save(mnnCorrected, file='Data/mnnCorrected_feature_k20.RData')
#load('Data/mnnCorrected_k200.RData')
mnnCorrected
```

Return mnn pairs, evaluate if sufficent number (else adjust k)
```{r check-mnn-pairs}
metadata(mnnCorrected)$merge.info

colData(mnnCorrected)
```

# Joint analysis

```{r joint-TSNE-UMAP}
library(gridExtra)

verbose_labels <- c(pigBALsce$verbose_cluster, pigPBMCsce$verbose_cluster)
mnnCorrected$verbose_cluster <- verbose_labels 
colData(mnnCorrected)

set.seed(123)
mnnCorrected <- fixedPCA(mnnCorrected, subset.row=NULL, rank = 50, assay.type='corrected')
#mnnCorrected <- runTSNE(mnnCorrected, dimred='PCA', n_dimred = 1:10, perplexity=20)
mnnCorrected <- runUMAP(mnnCorrected, dimred='PCA', n_dimred = 1:20, n_neighbors=20)

png(filename='Results/Combined_mnnCorrectedUMAP.png', height=1080, width=2160)
plot1 <- plotUMAP(mnnCorrected, colour_by="verbose_cluster", point_size=0.3) +
  theme(legend.position="bottom", legend.direction="vertical") +
  guides(colour = guide_legend(override.aes = list(size=4)))+ theme(text = element_text(size = 10)) +
  theme(aspect.ratio = 1)
plot2 <- plotUMAP(mnnCorrected, colour_by="batch", point_size=0.3) +
  theme(legend.position="bottom", legend.direction="vertical") +
  guides(colour = guide_legend(override.aes = list(size=4)))+ theme(text = element_text(size = 10)) +
  theme(aspect.ratio = 1)
grid.arrange(plot1, plot2, ncol=2)
dev.off()
```


# Wrapping up

Save everything

```{r save-everything}
sessionInfo()
save(list=ls(), file='Data/mnnCorrectSave.RData') 
```